\name: Deterministic and Coherent Implicits

### 2018-11-27 jfp/src/Translation.tex

It follows immediately that deterministic resolution is also coherent,
which for our elaboration-based setting is formulated, following Biernacki and Polesiuk~\shortcite{dariusz}, in
terms of \emph{contextual equivalence}~\cite{morris}.
\begin{corollary}[Coherence]\label{lem:coherence}
Let $\tenv$ be a type environment, $\rulet$ be a type and $E_1$, $E_2$ be system F expressions.
Assume that $\unambig{}{\rulet}$ and $\forall ?\rulet_i\in\tenv$, $\unambig{}{\rulet_i}$. Then
\[ %\forall \tenv, \rulet, E_1, E_2: \quad\quad 
     \dres{\tenv}{\rulet}{E_1} ~\wedge~ 
     \dres{\tenv}{\rulet}{E_2} \quad\Rightarrow\quad 
     \fctx{\etrans{\tenv}}{E_1}{E_2}{\ttrans[\rulet]} \]
\end{corollary}
Here the contextual equivalence judgment $\fctx{\tenv}{E_1}{E_2}{T}$ captures
that $E_1$ and $E_2$ behave the same in any well-typed program context. We do
not give the formal definition here\footnote{A definition can be found at \url{https://www.seas.harvard.edu/courses/cs252/2017fa/lectures/lec02-contextual-equiv.pdf}} as it is not required to prove the
corollary; we only need to know that it is an equivalence relation, and, more
specifically, that it is reflexive, i.e., any well-typed System~F term is
contextually equivalent to itself. 
% \bruno{Should we not mention that contextual equivalence for System F
%   is well-studied, thus interested readers can consult the definition
%   in other works?}

>>>>

It follows immediately that deterministic resolution is also coherent,
which for our elaboration-based setting is formulated, following Biernacki and Polesiuk~\shortcite{dariusz}, in
terms of \emph{contextual equivalence}~\cite{morris}.
\begin{corollary}[Coherence]\label{lem:coherence}
Let $\tenv$ be a type environment, $\rulet$ be a type and $E_1$, $E_2$ be system F expressions.
Assume that $\unambig{}{\rulet}$ and $\forall ?\rulet_i\in\tenv$, $\unambig{}{\rulet_i}$. Then
\[ %\forall \tenv, \rulet, E_1, E_2: \quad\quad 
     \dres{\tenv}{\rulet}{E_1} ~\wedge~ 
     \dres{\tenv}{\rulet}{E_2} \quad\Rightarrow\quad 
     \fctx{\etrans{\tenv}}{E_1}{E_2}{\ttrans[\rulet]} \]
\end{corollary}
Here, the well-studied~\cite{contextual_equivalence} contextual equivalence judgment $\fctx{\tenv}{E_1}{E_2}{T}$ captures
that $E_1$ and $E_2$ behave the same in any well-typed program context. We do
not give the formal definition here as it is not required to prove the
%\footnote{A definition can be found at \url{https://www.seas.harvard.edu/courses/cs252/2017fa/lectures/lec02-contextual-equiv.pdf}} 
corollary; we only need to know that it is an equivalence relation, and, more
specifically, that it is reflexive, i.e., any well-typed System~F term is
contextually equivalent to itself.
% \bruno{Should we not mention that contextual equivalence for System F
%   is well-studied, thus interested readers can consult the definition
%   in other works?}
